GPU Efficient Texture Based Bezier Curve Evaluation

### 2015-07-15 Paper/paper.tex
More progress, starting to run low on todos! perf next?

\teaser{
  \begin{tikzpicture}[x=5in,y=1.7in]
    \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=5in,height=1.7in]{Teaser.png}};
    % representative line for where the quadratic bezier curve lives in the bilinear sampled texture
    \draw[yellow,->,line width=0.5mm]    (0.41,0.25) -- (0.58,0.75);
    \draw (0.08,0.25) node[] {\Huge{A}};
    \draw (0.25,0.25) node[] {\Huge{B}};
    \draw (0.08,0.75) node[] {\Huge{B}};
    \draw (0.25,0.75) node[] {\Huge{C}};    
  \end{tikzpicture}
  \caption{\textit{Left:} 2x2 texture containing control points for a quadratic Bezier curve in each color channel. \textit{Middle:} The texture as viewed with bilinear sampling. \textit{Right:} The resulting curves when sampling along the yellow line (Alpha curve omitted).}
  \label{fig:teaser}
}

@@

A limitation with this technique is that control points can only take values that can be stored and recalled from the specific texture format you are using.  This can be an issue not only for the range of values possibly - such as when a texture can only hold values from 0 to 1 or 0 to 255 - but is also an issue for accuracy due being interpolated in a 9 bit (1.8) fixed point format \cite{NVIDIA}.   The value range problem is easily solved by scaling and offsetting.  The accuracy problem is not as easily addressed, but there are some ways mentioned later in the paper.

The accuracy issues can be seen in \autoref{fig:quickaccuracy}, although the accuracy is still significantly higher than a baked out curve using the same number of pixels as seen in \autoref{fig:quickcomparison}.

TODO: the 9 bit fixed point format doesn't belong mentioned where it is.
TODO: also need to find and mention the thing about how interpolation weights are multiplied before taking it into floating point.

@@

@@

A naive way to do this would be to use nearest neighbor texture sampling instead of linear, and do the linear interpolations yourself in the shader program.  This would mean twice as many texture reads with a 1D texture, four times as many texture reads with a 2D texture, and eight times as many texture reads with a 3D texture!  You would also be doing a lot more blending in the shader program, so your shader programs would become heavier weight.  This solution is essentially doing software linear texture sampling.

@@

For instance, if you had a 2x2 image which encoded a quadratic Bezier curve, you might do two texture reads, letting the texture sampler do the $x$ axis interpolation, and then you could linearly interpolate those results in the shader program in full precision floating point math.  This would give you a result somewhere between full hardware and full software interpolation.

TODO: image to show hw vs sw vs hwsw

@@

@@

TODO: picture of halfspace and gradient distance functions.

@@

TODO: do we need an image of polar form? maybe combine with picture above and reference the same thing

@@

Lastly, in regards to vector graphics, this technique can be used to allow people to define custom color gradients.  The benefit is that custom, non linear color gradients can be defined, and can be scaled up without decaying to the usual artifacts seen from linear texture sampling.

TODO: do we need a gradient (rect / polar?) image?
TODO: need an image of the flag

%-------------------------------------------------------------------------
\section{Comparisons With Other Techniques}
\label{sec:comparisonswithothertechniques}

TODO: real apples to apples comparisons with texture baking (can be any order!) as well as shader constants.
TODO: maybe we don't need this

@@

TODO: Show some real performance values. 
TODO: show difference between 3d texture doing cubic, vs 2d texture with 2 texture reads and shader program results.
TODO: show results from multiple GPUs?

@@

There may also be usefulness in storing curves in polar form to aproximate lobes for lighting algorithms.

>>>>

\teaser{
  \begin{tikzpicture}[x=5in,y=1.7in]
    \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=5in,height=1.7in]{Teaser.png}};
    % representative line for where the quadratic bezier curve lives in the bilinear sampled texture
    \draw[yellow,->,line width=0.5mm]    (0.41,0.25) -- (0.58,0.75);
    \draw (0.08,0.25) node[] {\Huge{A}};
    \draw (0.25,0.25) node[] {\Huge{B}};
    \draw (0.08,0.75) node[] {\Huge{B}};
    \draw (0.25,0.75) node[] {\Huge{C}};    
  \end{tikzpicture}
  \caption{\textit{Left:} 2x2 texture containing control points for a quadratic Bezier curve in each color channel. \textit{Middle:} The texture as viewed with bilinear sampling. \textit{Right:} The resulting quadratic Bezier curves when sampling along the yellow line (Alpha curve omitted).}
  \label{fig:teaser}
}

@@

A limitation with this technique is that control points can only take values that can be stored and recalled from the specific texture format being used, which puts a limitation on the range of values able to be used, as well as the precision of the values stored.

Furthermore, the fractional position of a texel used in the interpolation formulas is limited to 256 values due to it being calculated with a 9 bit (1.8) fixed point format \cite{NVIDIA}.

The accuracy issues can be seen in \autoref{fig:quickaccuracy}, although the accuracy can still be significantly higher than a baked out curve using the same number of pixels as seen in \autoref{fig:quickcomparison}.

@@

  \begin{figure}
    \begin{tikzpicture}[x=12cm,y=4cm]
      % border
      \draw (0cm,0cm) rectangle (12cm,4cm);
      % images
      \node[anchor=south west,inner sep=0] at (0cm,0cm) {\includegraphics[width=4cm,height=4cm]{SampleHW.png}};
      \node[anchor=south west,inner sep=0] at (4cm,0cm) {\includegraphics[width=4cm,height=4cm]{SampleHWSW.png}};
      \node[anchor=south west,inner sep=0] at (8cm,0cm) {\includegraphics[width=4cm,height=4cm]{SampleSW.png}};

    \end{tikzpicture}
    \caption{\textit{Left:} Curve evaluted with a single texture read and bilinear texture sampling.  \textit{Middle:} Curve evaluated with two linear texture samples, combined in the shader program.  \textit{Right:} Curve evaluated with three texture samples (one per control point), combined in the shader program.}   
    \label{fig:samplehwsw}
  \end{figure}

@@

A naive way to do this would be to use nearest neighbor texture sampling instead of linear, and do the linear interpolations yourself in the shader program.  This would mean twice as many texture reads with a 1D texture, four times as many texture reads with a 2D texture, and eight times as many texture reads with a 3D texture.  You would also be doing a lot more blending in the shader program, so your shader programs would become heavier weight.  This solution is essentially doing software linear texture sampling.

@@

For instance, if you had a 2x2 image which encoded a quadratic Bezier curve, you might do two texture reads, letting the texture sampler do the $x$ axis interpolation, and then you could linearly interpolate those results in the shader program in full precision floating point math.  This would give you a result somewhere between full hardware and full software interpolation.  You can compare the results in \autoref{fig:samplehwsw}.

@@

  \begin{figure}
    \hspace{-1.5cm}
    \begin{tikzpicture}[x=16cm,y=4cm]
      % images
      \node[anchor=south west,inner sep=0] at (0cm,0cm) {\includegraphics[width=8cm,height=4cm]{Flag.png}};
      \node[anchor=south west,inner sep=0] at (8cm,0cm) {\includegraphics[width=8cm,height=4cm]{FlagZoom.png}};
    \end{tikzpicture}
    \caption{\textit{Left:}An animated flag textured with vector graphics using the techniques in this paper, using an 8x4 RGBA source texture to generate curves and color gradients.  \textit{Right:} Zooming in on part of the image to show the quality of the vector graphics.}   
    \label{fig:vectorgfx}
  \end{figure}

@@

@@

@@

Lastly, in regards to vector graphics, this technique can be used to allow people to define custom color gradients.  The benefit is that custom, non linear color gradients can be defined, and can be zoomed into without decaying to linear interpolation.

A demonstration of these usage cases can be seen in \autoref{fig:vectorgfx}.

@@

TODO: perf goals are...
* show perf of a few curve types (linear, quadratic, cubic, quartic?), using 1d, 2d, 3d textures.  show based on \# of texture reads (like hw, hwsw, sw).
* show results from laptop as well as work machine. maybe laptop on single GPU?

TODO: perf tests... at 1000x1000
* linear    :  1d texture (1 read vs 2 reads then lerp)
* quadratic :  1d texture (2 reads and lerp vs 3 reads and quadratic bezier), 2d texture (1 read vs 2 reads and lerp vs 3 reads and quadratic Bezier)
* cubic     :  1d texture (3 reads and quad vs 4 reads and quartic bezier), 2d texture (), 3d texture 
* quartic   :  1d texture, 2d texture, 3d texture

TODO: mention that it matters if you are texture lookup bound or not.  correct terms for that = ??

@@

There may also be usefulness in curves in polar form to aproximate lobes for lighting algorithms.
